---
title: "Machine Learning Model Building for Activity Tracking Device Data"
output:
  html_document:
    keep_md: yes
date: "Wednesday, May 20, 2015"
---
###Synopsis
This report applies the practical machine learning algorithm to fit the activity tracking device data of a group of people doing barbell lifts correctly and incorrectly in 5 different ways, builds a model in order to predict the manner in which they did the exercise.

###Data Processing
Load required packages and data:
```{r}
library(caret)

data <- read.csv("pml-training.csv")
```
Un-select first 6 columns of index, username and timestamp:
```{r}
data <- data[,7:160]

```
Un-select columns with mostly missing values.
```{r}
data[data=='']<-NA
naa <- apply(data,2,function(x) mean(is.na(x)))
data_clean <- data[,naa<0.5]
```

###Model Building
Split data into training set and testing set:
```{r}
InTrain <- createDataPartition(data_clean$classe,p=0.75,list=FALSE)
training <- data_clean[InTrain,]
testing <- data_clean[-InTrain,]
```

Set fit control parameters: using 5-fold cross validation
```{r}
fitControl <- trainControl(method="cv",number=5,repeats=2,allowParallel=TRUE)
tgrid <- expand.grid(mtry=c(6))
```
Fit a boosting tree model via gbm package:
```{r}
set.seed(565)
rfFit1 <- train(classe~., data=training,method='rf',trControl=fitControl,tuneGrid = tgrid, verbose=TRUE)
rfFit1
print(rfFit1$finalModel)
```
The out of sample error estimated by the cross-validation is <0.5%, which is very low. 

###Model Testing
Do confusionMatrix on the testing data set for out of sample error:
```{r}
cm <- confusionMatrix(testing$classe,predict(rfFit1,testing))
cm
```
The accuracy is >0.99, which is very high accuracy.
